Technical Report

### 2018-06-22 fully-abstract-comp/technical-report.tex
Comment out obsolete notes.

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \Caps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    \scbnf \in &\SealableCaps&\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    & & & {\sourcecolor{} \; \mid \retptrd(\basebnf,\aendbnf) \mid \retptrc(\basebnf,\aendbnf,\addrbnf)}\\
    \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}\\ 
    &\rbnf & \defbnf &\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Addr \times \MemSeg}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus \{\halted\}
  \end{array}
\]
%\lau{06-09-2017: Should seals also have the option to be linear?}
%\lau{13-09-2017: For now I have added linearity to the seals.}
%\lau{15-12-2017: Removed linearity from seal - I could not come up with a single good reason to have it as we have them here.}
\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
\lau{28-09-2017: Why would you like this? If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail. We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part. Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}. This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame. Makes sense?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\rbnf$ defines the finite set $\RegName$. 
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

In the source language, $\src{\Stack}$ is a call stack that contains the data for each call. The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc and 2) caller's private stack.
\lau{18-12-2017: Stack frame used to contain the address of the old stack pointer. This should be inferable from other information, so it was removed.}

@@

\noindent\textbf{call}\\
\lau{19-09-2017: How do we get the $\sigma_?$ for the return pointers? It should be in memory somewhere.}
\lau{19-09-2017: In which cases do we have undefined here?}
\lau{25-10-2017: I have forgotten why we save $e_\stk$ on the stack and whether this is still necessary (it at least prevent the empty case, but I do not recall if it is necessary for other reasons).}
\lau{20-12-2017: At some point, I added an offset annotation $o$ to all calls. I believe this removes the need for a function that annotate seals. It will still have to be checked that all calls in a program uses different offsets, but I suspect this can be done as some assumption in what we want to prove.}
\lau{20-12-2017: Until this day, we had a $\sealAss{}$ function: This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack. We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}}

\begin{multline*}
  \sem{\scall{\offpc,\offsigma}{\src{r_1}}{\src{r_2}}}(\Phi) = \\
  \begin{cases}
    \mathit{xjumpResult}\left(c_1,c_2, 
     \arraycolsep=0pt
    \array[]{rl}
      \Phi'&\updReg{r_1}{w_1} \\
      &\updReg{r_2}{w_2} \\
      &\updReg{\rstk}{c_\stk}\\
      &\updReg{\rretc}{\sealed{\sigma_?,c_\opc}}\\
      &\updReg{\rretd}{\sealed{\sigma_?,c_\var{priv\_data}}}\\
      &\updReg{\rtmp{1}}{0}
      \endarray
     \right)
    & \arraycolsep=0pt
    \begin{array}[t]{l}
      r_1 \neq \rtmp{1} \tand r_2 \neq \rtmp{1} \\
      \Phi(r_1) = \sealed{\sigma_1,c_1} \tand \\
      \Phi(r_2) = \sealed{\sigma_2,c_2} \tand \\
      \sigma_1 = \sigma_2 \tand \\
      \nonExec{c_2} \tand\\
      \Phi = (\mem,\reg,\stk,\ms_\stk) \tand\\
      \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \tand \\
      \baddr_\stk < \aaddr_\stk \leq \eaddr_\stk \tand \\
      \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} \tand\\
      \ms_{\stk,\var{rest}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} \tand \\
      c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
      c_\var{priv\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) \tand \\
      \Phi(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) \tand \\
      \opc = \aaddr + \calllen \tand \\
      c_\opc = \retptrc(\baddr,\eaddr,\aaddr+\calllen) \tand \\
      \stk' = (\opc,\ms_{\stk,\priv}) :: \stk \tand\\
      \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
      \mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}\tand\\
      \baddr \leq  \aaddr+\offpc \leq \eaddr \tand \\
      \sigma_? = \sigma_\aaddr + \offsigma \tand\\
      \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand\\
      w_1 = \linCons{\Phi(r_1)} \tand \\
      w_2 = \linCons{\Phi(r_2)}
    \end{array}\\
    \failed & \totherwise
  \end{cases}
\end{multline*}
\lau{20-12-2017: After looking this over, I have a couple of questions (for myself and everyone else): In the condition for call, we assume $\Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$. There is, however, no check in the implementation of call that ensures this. That says, we are going to have an assumption that the stack pointer base address has been checked before all calls (i.e., that it is $\stkb$) which would indeed mean that it is a stack pointer (assuming that the address $\stkb$ is a stack address to begin with). Why did we not want to check the stack end address here? And if we do not check it as part of the call implementation, do we then want to match the call implementation better?}
\dominique{13-3-2018: re the above note: the above should be mitigated by Definition~\ref{def:check-stack-addr-before-call}. Adding a check is in principle not necessary, since the check would only protect trusted code from itself, rather than the adversary.}
\dominique{13-3-2018: where does $o$ come from above in the calculation of $\sigma_?$?}

>>>>

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \Caps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    \scbnf \in &\SealableCaps&\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    & & & {\sourcecolor{} \; \mid \retptrd(\basebnf,\aendbnf) \mid \retptrc(\basebnf,\aendbnf,\addrbnf)}\\
    \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}\\ 
    &\rbnf & \defbnf &\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Addr \times \MemSeg}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus \{\halted\}
  \end{array}
\]
%\lau{06-09-2017: Should seals also have the option to be linear?}
%\lau{13-09-2017: For now I have added linearity to the seals.}
%\lau{15-12-2017: Removed linearity from seal - I could not come up with a single good reason to have it as we have them here.}
% \dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
% \lau{28-09-2017: Why would you like this? If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail. We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
% \dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
% \lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part. Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}. This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame. Makes sense?}
% \dominique{22-6-2018: yes}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\rbnf$ defines the finite set $\RegName$. 
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

In the source language, $\src{\Stack}$ is a call stack that contains the data for each call. The call stack consists of a number of $\src{\StkFrame}$'s that contains 1) the old pc and 2) caller's private stack.
% \lau{18-12-2017: Stack frame used to contain the address of the old stack pointer. This should be inferable from other information, so it was removed.}

@@

\noindent\textbf{call}\\
% \lau{19-09-2017: How do we get the $\sigma_?$ for the return pointers? It should be in memory somewhere.}
% \lau{19-09-2017: In which cases do we have undefined here?}
% \lau{25-10-2017: I have forgotten why we save $e_\stk$ on the stack and whether this is still necessary (it at least prevent the empty case, but I do not recall if it is necessary for other reasons).}
% \lau{20-12-2017: At some point, I added an offset annotation $o$ to all calls. I believe this removes the need for a function that annotate seals. It will still have to be checked that all calls in a program uses different offsets, but I suspect this can be done as some assumption in what we want to prove.}
% \lau{20-12-2017: Until this day, we had a $\sealAss{}$ function: This instruction assumes that the current address of the stack pointer, $\aaddr_\stk$, points to the first free address of the stack. We assume a seal assignment function $\sealAss{}$ that given an address assigns a seal.\lau{25-09-2017: This will need some more explanation - namely, there should be some relationship between the available seals and this function (I also need to think more about the specifics of this).}}

\begin{multline*}
  \sem{\scall{\offpc,\offsigma}{\src{r_1}}{\src{r_2}}}(\Phi) = \\
  \begin{cases}
    \mathit{xjumpResult}\left(c_1,c_2, 
     \arraycolsep=0pt
    \array[]{rl}
      \Phi'&\updReg{r_1}{w_1} \\
      &\updReg{r_2}{w_2} \\
      &\updReg{\rstk}{c_\stk}\\
      &\updReg{\rretc}{\sealed{\sigma_?,c_\opc}}\\
      &\updReg{\rretd}{\sealed{\sigma_?,c_\var{priv\_data}}}\\
      &\updReg{\rtmp{1}}{0}
      \endarray
     \right)
    & \arraycolsep=0pt
    \begin{array}[t]{l}
      r_1 \neq \rtmp{1} \tand r_2 \neq \rtmp{1} \\
      \Phi(r_1) = \sealed{\sigma_1,c_1} \tand \\
      \Phi(r_2) = \sealed{\sigma_2,c_2} \tand \\
      \sigma_1 = \sigma_2 \tand \\
      \nonExec{c_2} \tand\\
      \Phi = (\mem,\reg,\stk,\ms_\stk) \tand\\
      \Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk} \tand \\
      \baddr_\stk < \aaddr_\stk \leq \eaddr_\stk \tand \\
      \ms_{\stk,\priv} = \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]}\update{\aaddr_\stk}{42} \tand\\
      \ms_{\stk,\var{rest}} = \ms_\stk - \ms_\stk |_{[\aaddr_\stk,\eaddr_\stk]} \tand \\
      c_\stk = \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1} \tand \\
      c_\var{priv\_data} = \retptrd(\aaddr_\stk,\eaddr_\stk) \tand \\
      \Phi(\pcreg) = ((\_,\_),\baddr,\eaddr,\aaddr) \tand \\
      \opc = \aaddr + \calllen \tand \\
      c_\opc = \retptrc(\baddr,\eaddr,\aaddr+\calllen) \tand \\
      \stk' = (\opc,\ms_{\stk,\priv}) :: \stk \tand\\
      \Phi' = (\mem,\reg,\stk',\ms_{\stk,\var{rest}})\tand\\
      \mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}\tand\\
      \baddr \leq  \aaddr+\offpc \leq \eaddr \tand \\
      \sigma_? = \sigma_\aaddr + \offsigma \tand\\
      \sigma_\baddr \leq \sigma_? \leq \sigma_\eaddr \tand\\
      w_1 = \linCons{\Phi(r_1)} \tand \\
      w_2 = \linCons{\Phi(r_2)}
    \end{array}\\
    \failed & \totherwise
  \end{cases}
\end{multline*}
% \lau{20-12-2017: After looking this over, I have a couple of questions (for myself and everyone else): In the condition for call, we assume $\Phi(\rstk) = \stkptr{\rw,\baddr_\stk,\eaddr_\stk,\aaddr_\stk}$. There is, however, no check in the implementation of call that ensures this. That says, we are going to have an assumption that the stack pointer base address has been checked before all calls (i.e., that it is $\stkb$) which would indeed mean that it is a stack pointer (assuming that the address $\stkb$ is a stack address to begin with). Why did we not want to check the stack end address here? And if we do not check it as part of the call implementation, do we then want to match the call implementation better?}
% \dominique{13-3-2018: re the above note: the above should be mitigated by Definition~\ref{def:check-stack-addr-before-call}. Adding a check is in principle not necessary, since the check would only protect trusted code from itself, rather than the adversary.}
% \dominique{13-3-2018: where does $o$ come from above in the calculation of $\sigma_?$?}
